---
title: "Figure S3"
author: "Sergio Garcia-Moreno Alcantara (sga@stowers.org) "
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
  html_document:
    depth: 3
    theme: sandstone
    toc: yes
    toc_float: yes
    code_folding: "hide"
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{1028-01444-003-001}
- \fancyfoot[LE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: inline
---

# Aim

The aim of this analysis is to investigate whether there are differences in the occupancy of the SAGA core module at different promoter types in the ovaries

# Enviroment setup

Set working directory and load required libraries and lab functions

```{r setup, warning=F, message=F, tidy=TRUE}

setwd("/n/projects/sga/analysis/SAGA/saga_publication/")
options(knitr.figure_dir = "plots/figure_s3/")

#Standard packages
library(GenomicAlignments); library(GenomicRanges); library(Biostrings); library(BSgenome.Dmelanogaster.UCSC.dm6);
library(TxDb.Dmelanogaster.UCSC.dm6.ensGene); library(dplyr); library(reshape2); library(plyranges); library(CAGEr); 
library(magrittr); library(ggplot2); library(cowplot); library(ggseqlogo); library(gridExtra); library(ggpubr)


#Lab sources
source("./shared_code/granges_common.r")
source("./shared_code/metapeak_common.r")
source("./shared_code/knitr_common.r")

```

# Analysis

## 1. Loading samples and necessary data sets

```{r load samples, tidy=TRUE}

sample_list <- read.csv("./chipseq_samples.csv", sep=";")

load_bigwig <- function(sample_list){

          bw_path <- function(path) {
     path = path
            }
     sample_list %>%
     mutate(list = purrr::map(as.character(path), bw_path)
     ) %>%  pull(list)
 }

bw_list <- load_bigwig(sample_list)
names(bw_list) <- sample_list$short_name

tss <- get(load("./rdata/dm6_mrna_ovaries_tss.RData"))

motif_list_ovaries <- get(load("./rdata/motif_list_ovaries.RData"))

```

## 2. Generate a metapeak plot for each SAGA core subunits at the different promoter types

```{r metapeak, warning=F, message=F, fig.height=8, fig.width=8, tidy=TRUE}

## Calculate the average signal per factor per base pair at different promoter types

promoter_type_metapeak_df <- mclapply(names(motif_list_ovaries), function(x){
  motif <- motif_list_ovaries[[x]]
  mclapply(names(bw_list), function(y){
  bw <- bw_list[[y]]
  standard_metapeak(motif, bw, 1000, 1001, paste(x,"at",y), NA)
}, mc.cores = 10)
  }, mc.cores = 10) %>% 
  do.call(c, .) %>% 
  bind_rows()

promoter_type_metapeak_df %<>% 
  mutate(., factor=gsub(".* ", "", promoter_type_metapeak_df$sample_name)) %<>% 
  mutate(., motif=gsub(" .*", "", promoter_type_metapeak_df$sample_name))

## Setting the plotting order

sample_levels <- c("ada2b", "spt3",  "saf6", "wda", "pol2")
motif_levels <- c("tata", "dpe", "tct", "hk")

## Create a plotting function

plot_func <- function(df, name, color){
          
            df$motif <- factor(df$motif, levels = c(motif_levels))
                  ggplot(df, aes(x=tss_distance, y=reads)) + 
                   geom_area(aes(fill=factor), alpha=0.6, show.legend = F) + 
                    scale_fill_manual(values = color) +
                    geom_vline(xintercept=0, linetype=2) +
                    facet_grid(factor~motif, scales = "free") +
                      ggtitle(name) + 
                      theme_cowplot() +
                      theme(plot.title = element_text(size = 15, face = "bold")) +
                      xlab("Distance from TSS (bp)") + ylab("Average RPM")
}

plot_func(promoter_type_metapeak_df, "SAGA metapeaks", c("#C62606", "#6D0068", "#0B2E9B", "#044F2A", "#AB2A35"))

```

## 3. Plot the signal distribution of each SAGA core subunit at the different promoter types

Signal was calculated as max height of the peaks for each promoter in 500 bp window centered at the TSS

```{r boxplot, warning=F, message=F, fig.height=8, fig.width=3, tidy=TRUE}

#Make a data frame containing transcript ID and total signal per gene and promoter type

sig_df <- mclapply(names(motif_list_ovaries), function(x){
                      
                      motif <- motif_list_ovaries[[x]]
                      mclapply(names(bw_list), function(y){
                      bw <- bw_list[[y]]
                
                      df <- data.frame(fb_t_id = motif$fb_t_id, 
                                       signal = regionMaxs(resize(motif, 501, "center"), bw),
                                       sample = y,
                                       motif = x)
                      df
                      }, mc.cores = 5)
                  }, mc.cores = 5) %>% do.call(c, .) %>% bind_rows()

sig_df$motif <- factor(sig_df$motif, levels = motif_levels)
sig_df$sample <- factor(sig_df$sample, levels = sample_levels)

sig_df <- filter(sig_df,signal>=0)

boxplot <- ggplot(sig_df,  aes(motif, log2(signal+1), fill=motif)) + 
  geom_boxplot(alpha=0.7)+
  theme_cowplot() +
  scale_fill_manual(values = c("indianred3", "#EE962B", "#228232", "#2C699B")) +
  ggtitle("SAGA total signal at promoter types",) +
  facet_wrap(~sample, scales = "free_y", ncol=1) +
  theme(plot.title = element_text(size = 15, face = "bold")) +
  xlab("Distance to TSS (bp)") + ylab("RPM (log2)")

boxplot

```

# Session Info

```{r session}
sessionInfo()
```
